---
title: "Lecture_2_Code_Example"
author: "Cheng"
output: pdf_document
---

# Lecture 2 R Session II
Content: 
1. object types: vector, matrix\*, array\*, factor, data frame, list
2. operators\*
3. functions: bult-in\*, user-defined
4. packages

# 1. Object types(array, data frame, list)

## 1.1 vector

Create vector

```{r}
v1 <- c(1, 2, 3)
v2 <- 1:10
v3 <- seq(1, 10, by = 2)
v.combind <- c(v1, v2)
print(v.combind)
```

Access element in vector

```{r}
# Accessing vector elements using position.
week.day <- c("Sun","Mon","Tue","Wed","Thurs","Fri","Sat")
week.day.subset <- week.day[c(2, 3, 6)]
print(week.day.subset)

# Accessing vector elements using logical indexing.
week.day.subset <- week.day[c(TRUE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE)]
print(week.day.subset)

# Accessing vector elements using 0/1 indexing.
week.day.subset <- week.day[c(0, 0, 0, 0, 0, 0, 1)]
print(week.day.subset)

# Accessing vector elements using negative indexing.
week.day.subset <- week.day[c(-1)]
print(week.day.subset)

# Replace an element in a vector 
week.day.subset[1] <- 1
print(week.day.subset)

# Add elements to vector
week.day.subset <- append(week.day.subset, 6)
print(week.day.subset)
```

Calculation of vector (element-wise)

Be careful of the vector lenght.

```{r}
# Create two vectors
v1 <- c(1, 2, 3, 4, 5)
v2 <- c(5:1)

# Vector addition
add.result <- v1 + v2
print(add.result)

# Vector subtraction
sub.result <- v1 - v2
print(sub.result)

# Vector multiplication
multi.result <- v1 * v2
print(multi.result)

# Vector division.
divi.result <- v1 / v2
print(divi.result)

# How about inner product?
#inner.product <- 
#print(inner.product)
```

## 1.2* Matrix

Create matrix

```{r}
# Elements are arranged sequentially by row.
M <- matrix(c(1:9), nrow = 3, byrow = TRUE)
print(M)

# Define the column and row names.
rownames = c("row1", "row2", "row3")
colnames = c("col1", "col2", "col3")

P <- matrix(c(1:9), nrow = 3, byrow = TRUE, dimnames = list(rownames, colnames))
print(P)
```

Access elements in matrix

```{r}
# Access the element at 2nd column and 1st row.
print(P[1,2])

# Access part of the matrix
print(P[1:2, 1:2])

# Access only the  2nd row.
print(P[2,])
```

Matrix calculation 

```{r}
# Create two 2x3 matrices.
matrix1 <- matrix(c(1: 4), nrow = 2)
print(matrix1)

matrix2 <- matrix(c(4: 1), nrow = 2)
print(matrix2)

# Add the matrices.
result <- matrix1 + matrix2
print(result)

# Subtract the matrices
result <- matrix1 - matrix2
print(result)

# Multiply the matrices.
result <- matrix1 * matrix2
print(result)

# Divide the matrices
result <- matrix1 / matrix2
print(result)

# Inner product
result <- matrix1 %*% matrix2
print(result)
```

## 1.3* array: 

array is a vector with one or more dimensions

```{r array}
# create an array
a1 <- array(data = c(1,2,3), dim = c(3), dimnames = NULL)# Create an one dimension array
a1 <- array(1,2,3) # wrong! take vector or list as input

# high dimensional array
a2 <- array(data=0,dim = c(2,3))# create a 2*3 matrix. two dimension array is almost the same as a matrix
a3 <- array(0, c(2,3)) # simplified syntax
a4 <- array(dim=c(3,3)) # create an empty matrix

# practice after class
vector1 <- c(1,2,3)
vector2 <- c(4,5,6,7,8,9)
a3 <- array(data = c(vector1, vector2),dim = c(3,3)) # add two vectors into an array. pay attention to the order of numbers
print (a3)
a3 <- array(c(vector1, vector2),dim = c(3,3,4)) # add two vectors into an array. pay attention to the order of numbers
print (a3)
```

array cannot contain different data types

```{r}
a <- 2
b <- "world"
a2 <- array(c(a,b))# all non-character elements would be transferred to character
c <- T
a2 <- array(c(a,c))# logical elements would be tranferred to number
```

elements in array(indexing)

```{r}
print(a3[1,3,2])# print the element in first row and third column of the second matrix
print(a3[1,,2])# print the first row of the second matrix

# replace elements in array
a4[,2:3] = vector2
a4
```

name columns and rows

```{r}
# by initialization
colName = c("c1","c2","c3") 
rowName = c("r1","r2","r3")
a4 = array(c(vector1, vector2), dim = c(3,3,1), dimnames = list(rowName,colName))

# reset name
colName = c("c01","c02","c03")
rowName = c("r01", "r02", "r03")
rownames(a4) = rowName# Retrieve or set the row names of a matrix-like object.
colnames(a4) = colName

a4["r01","c02"]
```

## 1.4 factor: 

Factor is used to categorize the data and store it as levels.
They can store both strings and integers.
they are useful in the columns which have a limited number of unique values. 
Like "Male, "Female" and True, False

```{r factor}
# encode vector as factor
data <- c("East","West","East","North","South","North","East","West","West","West","East","North")
print(data)
print(is.factor(data))

# Apply the factor function.
factor_data <- factor(data)
print(is.factor(factor_data))
print(factor_data)

# Change element values
factor_data <- factor(data, labels = c(1, 2, 3, 4))
print(factor_data)
```

Easy to store and compare

```{r}
# store as integer
month.vec <- c("March","April","January","November","January", "September","October","September","November","August",
          "January","November","November","February","May","August",
           "July","December","August","August","September","November",
           "February","April")
month.factor <- factor(month.vec)
table(month.factor)

# easy to compare strings
month.order = c("January","February","March",
               "April","May","June","July","August","September",
               "October","November","December")
month.factor.ordered <- factor(month.vec, levels = month.order, order = T)
month.factor.ordered[2] > month.factor.ordered[1]
print(month.factor.ordered)
```

## 1.5 dataframe:

A data frame is a table or a two-dimensional array-like structure in which each 
column contains values of one variable and each row contains one set of values 
from each column. Each column holds the same data type but not necessary among 
columns.It's a fundamental data structure by most of R's Packages.

```{r dataframe}
a <- c(1,2,3) # one observation can be regarded as one column
b <- c("beta","gamma","alpha")
df1 <- data.frame(a,b) # create a data frame
print(df1)

# second method (initialize with column name)
a <- c("Tom","Bob","Amy")# name columns
b <- c(20,53,12)
df2 <- data.frame(Name = a, Age = b)
print(df2)
```

access data frame element

```{r}
# first method([ ])
df2[2, 2] # get the second column
df2[2, 'Age']
df2['Age'][2,] # return data.frame
#print(class(df2['Name'])) # class(): type of object
#print(class(df2[,'Name']))

# second method( $ )
df2$Age# access one observation by name
df2$Name[-2] # get all the elements except the second one
```

Expand Data Frame:
add row and column to existing data frame

```{r}
# first method(rbind, cbind)
df2 <- cbind(df2, Male=c(T,T,F)) # Add a new column
print(df2)

df2 <- rbind(df2, data.frame(Name="Tom" ,Age=34, Male=T)) # Add a new row using vector
print(df2)

# second method(assiging values)
df2['col3'] <- c(1,2,3,4)
print(df2)

df2$new.col <- c(10, 20, 5, 2)
print(df2)
```

## 1.6 list 
List can hold items of different data types

```{r list}
# create a simple list
list1 <- list(df1, df2, a3)
list1

# named list
list2 = list(x = c(1, 2, 3), y = c(1, 4, 9))
list2
```

elements in list

```{r}
# [] vs [[]]
list2[1]
list2[[1]]
class(list2[1])# sublist of a list is also a list. Attention: won't get number 3
class(list2[[1]]) # access values of a list
list2[1][1] # return type = num
list2[[1]][1]
```

an good example: banking system *

```{r}
bankName = c("BOA", "Chase", "JPMorgan") 
bankEquity = c(210,230,304)
bankLinks = matrix(c(1,1,1,0,1,1,0,0,1), nrow = 3) 
bankLinks
bankInfo = list(bankName, bankEquity, bankLinks)  # add elements in list.
names(bankInfo) = c("Names", "Equities", "Links") # give names to the elements in a list
print(bankInfo$Names) # display with name. equals to print (x[1])
```

#2*. Operations

## 2.1* arithmetic operators(+-*/)

```{r operations}
# number case
num1 = 3 # operations of two data
num2 = 4
num1 + num2  # 
num2 %% num1 # modulo operation
# modulo division: (5 mod 2) =1, because 5/2 has a quotient of 2 and a remainder of 1

# vector case
v1 = c(1,2,3,4) # operators in a vector
v1 + 5 # every element in vector plus 5
v2 = c(2,2,2,2) # operations of two vectors
v1 + v2

# matrix case
mat1 = matrix(c(v1,v2), nrow = 2, byrow = FALSE)
mat2 = matrix(c(v1,v2), nrow = 2, byrow = TRUE)
mat1 * mat2
mat1 %*% t(mat2)
```

## 2.2* comparison operators(<, <= >= == !=)

```{r}
num1 < num2
num1 == num2 # attention: "a==b" is different from "a=b"
num1 != num2 # num1 NOT equal to num2

v = c(2, 5.5, 6, 9)
t = c(8, 2.5, 14, 9)
print(v > t)
```

## 2.3* logical operators(! & |)

```{r}
a = 3
a<5 & a>1
a<5 && a>1 # & and && indicat logical and
1<a<5 # wrong. it's necessary to use & 
a<5 & a<1
a<5 | a<1
a<5 || a<1 # | and || indicate logical or

3 & 4# attention: different from 3+4
```

# 3. Functions

## 3.1* built-in functions for numbers

```{r functions}
a <- 4 
sqrt(a) # square root
exp(a) # exoponential 
log(a) # logarithm
```

## 3.2* built-in functions for vectors

```{r}
c <- c(1,2,3,4,5)
r <- rnorm(100)
r

sqrt(c) # square root
log(c) # logrithm
min(c) # minimum number in a vector
min(c,d)
max(c) # maximum number
var(c) # variance
sd(c) # standard deviation
median(c) #median number
mean(c)
hist(r)
quantile(r)

summary(r) # descripe the data set using statistics
sort(c) # sort a vector/array
sort(c, decreasing = TRUE)
order(c)# order number of elements in a vector/array
union(x, y)
intersect(x, y)
setdiff(x, y)
setequal(x, y) # common elements in different vectors
```

## 3.3* built-in functions for matrix

```{r}
diag(m)# diagonal of a matrix

# Find the solution for Ab, where A=matrix(c(2,0,1,3), ncol=2), b=c(7,4).
A <- matrix(c(2,0,1,3), ncol=2)
b <- c(7,4)
b%*%A # Multiplies two matrices

# find the solution for AB, where B=matrix(c(5,2,4,-1), ncol=2)
B <- matrix(c(5,2,4,-1), ncol=2)
A%*%B

# Find the inverse matrix of A.
solve(A) 

# Find the value of x on Ax=b. (A^(-1) * b)
solve(A, b)# solves the equation A %*% x = b for x

# Using the function eigen find the eigenvalue for A.
# lamda * v = v * A
eigen(A)

# Find the eigenvalues and eigenvectors of A'A . Hint: Use crossprod to compute A'A .
eigen(crossprod(A))

t(A) %*% A == crossprod(A)
```

## 3.4 user-defined function

```{r}
# function without argument

myfunc1 <- function(){
  a <- c(1,2,3)
  a.square.root <- sqrt(a)
  return(a.square.root)
}

result <- myfunc1()
print(result)
myfunc1(c(1,2))# error: unused argument
```


```{r}
# function with arguments
myfunc2 <- function(arg1, arg2){
  a <- arg1 * arg2
  return(a)
}
c <- 3
result <- myfunc2(c,c) # call the new function
print(result)
myfunc2(c)# Error in myfunc2() : argument "arg1" is missing, with no default
```


```{r}
# function with default arguments
get.circle.area <- function(radius = 1, pi = 3.14){ 
  a <- pi * radius ^ 2
  return(a)
}

get.circle.area(radius=6)
get.circle.area(3, 3)
get.circle.area()
get.circle.area(5)# first argument is replaced by 5

```

3.4* other built-in functions(apply family)

apply(x,margin,function): Returns a vector or array or list of values obtained by applying a function to margins of an array or matrix. Margin is the name of the dataset used. 1 indicates rows, 2 indicates columns, 3 indicates matrix

```{r}
# create an example
M <- matrix(c(1:9), byrow = T, nrow = 3)
print(M)

# apply
result <- apply(X = M, MARGIN = 1, FUN = sum) # get sum of elements for matrix 
print(result)
```

'lapply' works with list (takes list as input and return a list)

```{r}
# take matrix as input return list
lapply(M, sum)

# takes list as input and return a list
x <- list(a = 1:10, beta = exp(-3:3), logic = c(TRUE,FALSE,FALSE,TRUE))
print(x)
lapply(x, mean)
```

'sapply' is a simple version (wrapper) of 'lapply' (work with list and return vector)

```{r}
# takes list as input and return a vector
sapply(x, mean)

# take matrix as input return list
sapply(M, sum)
```

# 4. Packages

```{r packages}
# 1. install a new package
install.packages(quantmod) # package 'quantmod' successfully unpacked and MD5 sums checked

# 2. load library
library("quantmod") 

# 3. call functions
Cl(AAPL)
getSymbols('AAPL')

quantmod::getSymbols()
A::getSymbols()

# 4. package management
library() # check list of all the packages installed
search() # get all packages currently loaded in the R environment
```

