---
title: "Lecture_3"
author: "Cheng"
output: pdf_document
---
# 0. Content: 
1. decision making
2. loop
3. input/output of data


# 1. Decision making
1.1 if
1.2 switch
1.3 verctorization

## 1.1 if else
```{r 1.1}
x <- c("what","is","truth")

print(x)

if("TRUTH" %in% x) {
   print("TRUTH is found")
} else {
   print("TRUTH is not found")
}


# else if 
x <- c("what","is","truth")

if("TRUTH" %in% x) {
   print("TRUTH is found the first time")
} else if ("truth" %in% x) {
   print("truth is found the second time")
} else {
   print("No truth found")
}
```

## 1.2 switch

Expanding 'if else' to switch

```{r}
x <- 'c'
if (x == "a") {
    "option 1"
  } else if (x == "b") {
    "option 2" 
  } else if (x == "c") {
    "option 3"
  } else {
    stop("Invalid `x` value")
  }
```

```{r}
x <- 'd'

switch(x,
      a = "option 1",
      b = "option 2",
      c = "option 3",
      print("Invalid `x` value")
)
```

------------ example 1 to practice ------------

```{r 1.2.1}
centre <- function(x, type) {
  switch(type,
         mean = mean(x),
         median = median(x),
         trimmed = mean(x, trim = 0.1))
}
x <- rcauchy(50)#  generate 50 random numbers based on the Cauchy distribution
centre(x, "mean")
centre(x, "median")
centre(x, "trimmed")
```

1.3 Vectorized if

Handling vectors of values is the job of ifelse():
a vectorised function with test, yes, and no vectors (that will be recycled to the same length):

```{r}
x <- 1:10

ifelse(x %% 2 == 0, "even", "odd")
#>  [1] "odd"  "even" "odd"  "even" "odd"  "even" "odd"  "even" "odd"  "even"

ifelse(x %% 5 == 0, "XXX", as.character(x))
#>  [1] "1"   "2"   "3"   "4"   "XXX" "6"   "7"   "8"   "9"   "XXX"
```

# 2. Loop
2.1 repeat loop
2.2 while loop
2.3 for loop
2.4 double loop
2.5 break loop
2.6 pipe

## 2.1 repeat loop
```{r 2.1}
a <- c("Hello")
num <- 1

repeat {
  print(a)
  num = num + 1
  if(num > 5){ break } # When the break is inside a loop, the loop is immediately terminated, and...
  # goes to the next statement following the loop
}
```

## 2.2 while loop
```{r 2.2}
# while loop
a <- c(1,23,4,5,5)
num <- 1

while (num <= length(a)) {
  print(a[num])
  num <- num + 1
}
```

repeat vs while
repeat loop checks the condition at the end of each iteration while the while loop checks it at 
the beginning of each iteration. So repeat loop executes at least one iteration but while 
loop may not execute any iterations if the condition is not fulfilled.

## 2.3 for loop
```{r 2.3}
# for loop
for (i in c('2', '3', 3, 4, 5)) {
  print(i)
}

for (i in seq(from = 1,to = 5, by = 0.1)){
  print (i/2)
}
```

## 2.4 double loop*

```{r 2.4}
# nesting for loop
matrix1 = matrix(c(1:9), nrow=3,ncol=3, byrow = T)
for(i in 1:3){
  for (j in 1:3){
    cat("row", i, 'column', j, 'value', matrix1[i, j], '\n')
  }
}
```

------------ example 2 to practice ------------

```{r}
# example: matrix dot product
dotProduct = function(m1,m2){ # matrix multiplication
  mat = matrix(nrow = dim(m1)[1], ncol = dim(m2)[2])
  for(i in 1:dim(m1)[1]){
    for (j in 1:dim(m2)[2]){
      mat[i,j] = sum(m1[i,]*m2[,j])
    }
  }
  return(mat)
}  
mat1 = matrix( 
     c(2, 4, 3, 1, 5, 7), # the data elements 
     nrow=2,              # number of rows 
     ncol=3,              # number of columns 
     byrow = TRUE)

mat2 = matrix( 
     c(2, 4, 3, 1, 5, 7), # the data elements 
     nrow=3,              # number of rows 
     ncol=2,              # number of columns 
     byrow = TRUE)

dotProduct(mat1, mat2)
```

## 2.5 break/next loop

break: When inside a loop, the loop is immediately terminated and program control resumes at 
the next statement following the loop.

```{r}
v <- LETTERS[1:6]

print(v)

for ( i in v) {
  if (i == "D") {
    break
  }
  print(i)
}
```

next: skips further evaluation and starts next iteration of the loop.

```{r}
v <- LETTERS[1:6]
for ( i in v) {
  if (i == "D") {
    next
  }
  print(i)
}
```

## 2.6 pipe

```{r}
x <- c(0.109, 0.359, 0.63, 0.996, 0.515, 0.142, 0.017, 0.829, 0.907)

# Compute the logarithm of `x`, return suitably lagged and iterated differences, 
# compute the exponential function and round the result
round(exp(diff(log(x), 1)), 1)

x.1 <- log(x)
x.2 <- diff(x.1)
x.3 <- exp(x.2)
x.final <- round(x.3)
```

Perform the same computations on `x` using pipe (%>%)

```{r}
library(magrittr)

x %>%
  log() %>%
  diff(1) %>%
  exp() %>%
  round(1)
```


# 3. I/O  data
3.1 RData file
3.2 text file
3.3 excel file
3.4 csv file
3.5* web file

## 3.1 RData file
read from RData
```{r}
path <- "/Users/cheng/Google Drive/PhD/Teaching/FE 513 Database Design/02_Codes/data/ETF.Rdata" # change the path accordingly
results <- load(path)
  
getwd()

path.desired <- '...'
setwd() <- path.desired
```

output RData

```{r}
path <- "data/ETF_copy.Rdata"
save(data, file = path)
```

## 3.2 text file
read a data table from txt file. 
the cells inside the table are separated by blank characters.

```{r}
path <- "data/AAPL.txt"
table.data <- read.table(path , fill = T, header = T)
head(table.data)
```

save data as a text file

```{r}
path <- "data/AAPL_copy.txt"
write.table(table.data, file = path)
```

## 3.3 excel file

Not recommend.
Try to use csv file to transfer data.

```{r}
library(readxl) # load package

path <- "data/mydata.xlsx"
myfile <- read_excel(path)
myfile
mysheet <- read_excel(path, sheet = "Sheet3")
mysheet
```

## 3.4 csv file 

read from csv file

```{r}
path <- "data/AAPL.csv"
aapl <- read.csv(file = path, head = TRUE) 
aapl
```

output data as a csv file

```{r}
path <- "data/AAPL_copy.csv"
write.csv(aapl, file = path)
```

## 3.5 web file*

```{r}
thepage <- readLines("https://stevensducks.com/schedule.aspx?path=mfence")
mypattern <- '<a aria-label=' #  search for matches to argument 
datalines <- grep(mypattern,thepage,value=TRUE) # get the lines where our data is. 
#grep() searches for matches to argument pattern within each element of a character vector
getString <- function(data){
  for (val in data){
    index1 = gregexpr(pattern = ">", val)[[1]][1]# get index of = in the string
    index2 = gregexpr(pattern = "<", val)[[1]][2]
    str = substring(val, index1+1, index2-1)
    print (str)
  }
}
getString(datalines)
# grep:search for matches to argument pattern within each element of a character vector
# gregexpr(pattern, str): find out position of pattern in str
str <- c(
  'Foo Bar Baz',
  'one two 3 four 5'
)

gregexpr("o", str)
gregexpr("\\s", str)# find spaces' positions
# regular expression: http://www.endmemo.com/program/R/gregexpr.php
```
